---
title: "Boston weather distributions"
author: "Gabriel Florit"
output:
  html_document:
    self_contained: false
---


```{r, message=F, warning=F, echo=F}

# set working directory
setwd("~/Documents/dev/weather-distributions")

# load libraries
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(weathermetrics)

```

```{r}

# read in data, parse date from string to date, add year column, convert max temp to fahrenheit
data <- read.csv('data/output/data.csv', stringsAsFactors = F) %>%
	mutate(date_ = ymd_hms(date_)) %>%
	mutate(year = year(date_)) %>%
	mutate(month = month(date_)) %>%
	mutate(tmax = celsius.to.fahrenheit(tmax))

```

***
#### What's the distribution of yearly measurement count?

```{r}

data %>%
	group_by(year) %>%
	summarise(count = n()) %>%
	ggplot(aes(count)) +
	geom_histogram(binwidth=1)

```

***
#### Looks like there are some years with less than 365 measurements. Which ones?

```{r}

data %>%
	group_by(year) %>%
	summarise(count = n()) %>%
	filter(count < 365)

```

How interesting - the 3 most recent years are missing data. Maybe Enigma hasn't updated their data? Either way we will only focus on complete years.

***
#### What's the overall distribution of max temperatures?

```{r}

complete <- data %>%
	filter(year < 2014)

complete %>%
	ggplot(aes(tmax)) +
	geom_histogram(binwidth=1) +
	labs(
		title='Distribution of Boston max temperatures',
		subtitle='1936-2013'
	)

```

Nice bell shape - looks mostly normal, slightly right-skewed.

***
#### What's the yearly median?

```{r, fig.height=2}

complete %>%
	group_by(year) %>%
	summarise(median = median(tmax)) %>%
	ggplot(aes(year, median)) +
	geom_line() +
	labs(
		title='Yearly median of Boston max temperatures',
		subtitle='1936-2013'
	)

```

This is all over the place. Doesn't say much.

***
#### What's the distribution of max temperature over time?

```{r, fig.height=2}

complete %>%
	ggplot(aes(x=year, y=tmax, group=year)) +
	geom_boxplot() +
	labs(
		title='Distribution of Boston max temperatures',
		subtitle='1936-2013'
	)

```

Also doesn't say much.

***
#### What about plotting density plots of all years?

```{r}

complete %>%
	ggplot(aes(x=tmax, group=year)) +
	geom_density()

```

Also doesn't say much. But looks pretty cool!

***
#### What's the distribution of max temperature over time, for July only?

```{r, fig.height=2}

complete %>%
	filter(month == 7) %>%
	ggplot(aes(x=year, y=tmax, group=year)) +
	geom_boxplot() +
	labs(
		title='Distribution of Boston max temperatures in July',
		subtitle='1936-2013'
	)

```
